% !TEX = slides.tex
\section{Addition of a New Constraint}
\label{sec:addition-new-constraint}

\begin{frame}{Addition of a New Constraint}
  \begin{itemize} \justifying \parskip4mm
    \item Addition of a new constraint may o may not affect the feasibility of the current optimal solution.
    \item It is sufficient to check whether new constraint is satified by the current optimal solution or not.
    \item If it is satisfied, the inclusion of the constraint has no effect on the current optimal solution \emph{i.e.,} it remains feasible as well as optimal. If, however, the constraint is not satified, the current \alert{optimal solution becomes infeasible. Dual simplex metho is then used to find the new optimal solution.}
  \end{itemize}
\end{frame}

\begin{frameExample}{\label{example:6.6-5.1}}{}
  % Example 6.6-5.1 gupta ebook
  \begin{onlyenv}<1>
    \begin{enumerate}[a)] \justifying \parskip4mm
      \item In example~\ref{example:6.6-2.1} an administrative constraint is added. Products $A, B$ and $C$ require 2, 3, and 2 hours of administrative services, while the total available administrative hours are 10. How does the optimal solution change?
      \item \label{example08} If the total available administrative time is 4 hours, find the new optimal solution.
    \end{enumerate}
  \end{onlyenv}


  \begin{onlyenv}<2>
    \[\max Z = 4x_1 +6x_2 +2x_3  \]

    {\centering
    subject to

    \vspace{3mm}
    \sysalign{r,r}%
    \sysdelim..%
    \systeme[x_1x_2x_3]%
    {
      x_1 + x_2 + x_3  \leq 3@(manpower),
      x_1 +4x_2 +7x_3 \leq 9@ (material),
      2x_1 +3x_2 +2x_3 \leq 10@ (administrative)
    }

    $x_1, x_2, x_3 \geq 0$
    \par}

    Where   $x_1, x_2, x_3$ are the number of products $A, B$ and $C$ produced.
  \end{onlyenv}

  \begin{onlyenv}<3>
    \[\max Z = 4x_1 +6x_2 +2x_3  \]

    {\centering
    subject to

    \vspace{3mm}
    \sysalign{r,r}%
    \sysdelim..%
    \systeme[x_1x_2x_3]%
    {
      x_1 + x_2 + x_3  \leq 3@(manpower),
      x_1 +4x_2 +7x_3 \leq 9@ (material),
      2x_1 +3x_2 +2x_3 \leq 4@ (administrative)
    }

    $x_1, x_2, x_3 \geq 0$
    \par}

    Where   $x_1, x_2, x_3$ are the number of products $A, B$ and $C$ produced.
  \end{onlyenv}
\end{frameExample}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "slides_sensitivity"
%%% End:
